Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-13 ms.tex
Added more references to previous classification papers

% Application of constraining mass profiles - large numbers needed.
The application of strong gravitational lensing to constrain the mass
distribution of strong lenses, such as early-type galaxies (ETGs),
necessitates large samples of galaxy-galaxy lensing systems.  The
strong lensing signature directly probes the underlying matter. These
systems exhibit a background source galaxy that the lens galaxy
deflects into a partial of full arc shaped Einstein ring.

@@

% Early efforts
Earlier efforts on automated, or ``robot'', identification of strong
lensing systems have two distinct steps.  The first enhances and
extracts characteristic features, and the second uses some form of
pattern recognition to classify lens and non-lens containing systems.
{\em ArcFinder} is one such code that finds arcs in groups or cluster
scale lens \citep{seidelandbartelmann_07}, and {\em RingFinder} an
analogous tool in searching for multiply imaged quasars
\citep{gavazzi_etal14}.

@@

% Describe how machine learning is a step up
With ``machine learning'' algorithms, we can train a model to separate
a dataset according to some desired classification scheme.  However,
we still need to reduce the dimensionality of the problem by removing
information in the images that do not pertain to the classifier.
Examples of automated classification will...

% Previous works and our goal to make a universal pipeline 
Machine learning to automate.  Previous works \citep{lenzen_etal04},
(StrongML), More+... Here, we focus on the improvement from
preprocessing the images and describe the pipeline, which is publicly
available.  Depending on the image type, different preprocessing is
likely necessary (rings, multiply lensed quasars, etc.) Our goal is to
create an open source pipeline that is general enough for the user to
select or add appropriate image processing techniques to augment the
features of the lensed image, train and test data with different
machine learning techniques, and to select the sequence of methods
that maximizes completeness with optimal purity.

% This is the first in MachLensFinder series with initial pipeline
% results
This is the first in a series of {\em MachLensFinder} papers.  We
present our image processing and machine learning classifier pipeline.
Our image processing procedure parameterizes edges in our image in a
1-d histogram, enabling fast feature extraction that is similar to
techniques used to identify humans in security software systems.  We
train a model with logistic regression to identify features
corresponding to strong lensing systems.

% Summary of results
We show results of the pipeline on mock galaxy-galaxy lens systems
observed by HST and LSST as respective examples of classifier
performance on optical space- and ground-based observations.
\citet{miraldaescudeandlehar_92} first suggested that massive
ellipticals would be likely frequent strong lensing sources in optical
surveys.  We train and test our pipeline on subsamples from 10,000
mock observed brightest cluster galaxy (BCG) strong lensing systems
and 10,000 BCG non-strong lensing systems.

% Need to optimize for completeness and purity.  Arc counts and mass
% reconstruction.
In general, lensed image finding must optimize for completeness and
purity.  In the application of arc abundance comparisons with
cosmological predictions, completeness is important; we must be able
to understand the reasons and frequency of false negatives.  In the
application of mass reconstruction of the lensing object, completeness
in strong lensing system identifications will allow us to constrain
the mass-to-light ratio for a large number of objects.

@@

%  Relevance of our work
While there has been a recent surge in the use of neural networks
applied to image classification problems, \wording{these
  implementations are not readily scalable, nor are the necessary
  computational resources and expertise easily accessible to the
  entire scientific community.}  We present a supervised
classification pipeline built with open source tools, tested on mock
HST and LSST data.  Results are reproducible on personal computers, as
both the pipeline and the data will be publicly distributed.  This
first paper serves as a case study for one feature extraction and
classical machine learning method, respectively, the {\em histogram of
  oriented gradients} and {\em logistic regression}.  This framework
is easily extendable to other feature methods and machine learning
algorithms for supervised classification.

>>>>

% Application of constraining mass profiles - large numbers needed.
The application of strong gravitational lensing to constrain the mass
distribution of strong lenses, such as early-type galaxies (ETGs),
necessitates large samples of galaxy-galaxy lensing systems.
\citet{miraldaescudeandlehar_92} first suggested that massive
ellipticals would be likely frequent strong lensing sources in optical
surveys.  These systems exhibit a background source galaxy that the
lens galaxy deflects into a partial of full arc shaped Einstein ring.
The strong lensing signature directly probes the underlying matter.
The identification of such systems in upcoming surveys is the first
step in constraining the mass-to-light ratio for a large number of
objects in this mass range.

@@

% Early efforts
Earlier efforts on automated, or ``robot'', identification of strong
lensing systems have two distinct generalized steps.  The first
enhances and extracts characteristic features, and the second uses
some form of pattern recognition in the features to classify lens and
non-lens containing systems.  Among others, selected features might
include shape parameterization
\citep[see][e.g.]{,alard_06,kuboanddellantonio_08,xu_etal16,lee_17},
colors in multi-band imaging \citep{maturi_etal14,gavazzi_etal14},
light profiles \citep{braultandgavazzi_15}, and characteristics of
potential lens galaxies \citep{marshall_etal09}.  Pattern recognition
often incorporate cutoffs in selected parameter space distributions
\citep{lenzen_etal04,joseph_etal14,paraficz_etal16}.  A number of
these have been publicly distributed, with specific end applications.
For example, {\em ArcFinder} is one such code that finds arcs in
groups or cluster scale lens \citep{seidelandbartelmann_07}, and {\em
  RingFinder} an analogous tool in searching for multiply imaged
quasars \citep{gavazzi_etal14}.  Codes like these have been
complementary to human identification \citep{more_etal16}.

% Classical machine learning methods
More recently, these methods have transitioned to using ``machine
learning'' algorithms.  In classical machine learning, we can train a
model to separate a dataset according to some desired classification
scheme.  However, we still need to reduce the dimensionality of the
problem by removing information in the images that do not pertain to
the classifier.

Examples of automated classification will...
% Previous works and our goal to make a universal pipeline 
Machine learning to automate.  Here, we focus on the improvement from
preprocessing the images and describe the pipeline, which is publicly
available.  Depending on the image type, different preprocessing is
likely necessary (rings, multiply lensed quasars, etc.)

@@

\citep{petrillo_etal17} convolutional neural networks with the Kilo
Degree Survey. \citet{bom_etal16} ... \citet{lanusse_etal17} using
state-of-the-art residual networks to....  directly in the image space
with minimal image pre-processing.

% Need to optimize for completeness and purity.  Arc counts and mass
% reconstruction.
In general, lensed image finding must optimize for completeness and
purity.  In the application of arc abundance comparisons with
cosmological predictions, completeness is important; we must be able
to understand the reasons and frequency of false negatives.  In the
application of mass reconstruction of the lensing object, completeness
in strong lensing system identifications will allow us

@@

%  Relevance of our work
While there has been a recent surge in the use of neural networks
applied to image classification problems, \wording{these
  implementations are not readily scalable, nor are the necessary
  computational resources and expertise easily accessible to the
  entire scientific community.}  We present a supervised
classification pipeline built with open source tools, tested on mock
HST and LSST data, which will also be made publicly available.  Our
goal is to centralize an open source pipeline that is general enough
for the user to select or add appropriate image processing techniques
to augment the features of the lensed image, train and test data with
different machine learning techniques, and to select the sequence of
methods that provides optimal model performance.  Results are
reproducible on personal computers, as both the pipeline and the data
will be publicly distributed.

% Summary of paper 
This first paper serves as a case study for our pipeline with one
feature extraction and classical machine learning method,
respectively, the {\em histogram of oriented gradients} (HOG) and {\em
  logistic regression} (LR).  HOG is a fast feature extraction
procedure that quantifies edges in images, commonly used to identify
humans in security software.  LR is a linear classifier, making its
scalability relatively straightforward with existing open source
tools.  This framework is easily extendable to other feature methods
and machine learning algorithms for supervised classification.

% Summary of results
We show results of the pipeline on mock galaxy-galaxy lens systems
observed by HST and LSST as respective examples of classifier
performance on optical space- and ground-based observations.  We train
and test our pipeline on subsamples from 10,000 mock observed strong
lensing systems and 10,000 non-strong lensing systems, each centered
on a potential lens galaxy.
